---
title: "If a Hoosier drops a pass in the end zone, but there's no one there to see it, did it ever really happen?"
output: html_document
---
```{r setup, include=FALSE}
# General set-up for the report.
# I find this useful, but you may improve upon, ignore, or remove
#  (possibly at your own peril).
# Tasks accomplished:
# Don't print out code
# Save results so that code blocks aren't re-run unless code
# changes (cache), _or_ a relevant earlier code block changed (autodep),
# don't clutter R output with messages or warnings (message, warning)
library(MASS)
library(knitr)
library(tidyverse)
library(dplyr)
library(ggthemes)
library(gridExtra)
library(ggplot2)
opts_chunk$set(echo=FALSE,
               cache=TRUE, autodep=TRUE,
               message=FALSE, warning=FALSE)
# Turn off meaningless clutter in summary() output
options(show.signif.stars=FALSE)
```

# By: Jordan Majoros, Minwoo Lee, Lester Blumberg, and Joseph Spatz

We are interested in finding out what motivates a person to dedicate a large chunk of their Saturday to attending a college football game. The conclusions that can be learned by analyzing this dataset have uses beyond just predicting future college football attendance. They can also be used, with much care, to understand the decisions people make when attending other events, athletic and non-athletic. Another reason why this dataset is interesting to our team is that we can relate our findings to behavior we have observed in our own college experiences.

# I. Introduction

```{r load-data, include=FALSE}
attendance.data = read.csv("CFBeattendance.csv")
newtime = str_split(attendance.data$Time, ":")
```

The data analyzed contains information primarily from NCAA FBS (Football Bowl Subdivision), and its Division I-A predecessor, College Football games spanning 2000 to 2018. The data is specifically the home games for the selected teams, with the corresponding attendance rates for the stadium played in. Some of the teams in the dataset have seasons included before they had reached the FBS level. For classification purposes, we list all of these teams' conferences as being FCS for those seasons. There are `r length(levels(attendance.data$Team))` teams recorded, totaling 6672 home games. The columns selected include date of the game, the home team, the opponent, time of game, rank of both teams, whether the game was televised, the result of the game, stadium capacity and fill rate, whether there is a new coach, as well as the weather metrics during the game (precipitation, snowfall, snow depth, temperature maximum, and temperature minimum). 

We are trying to determine the relationships between the variables and attendance? We’re looking at the relationships with attendance over the course of one season as well as over the many seasons in the sample. Have these relationships changed over time? For example, are there some predictors that had a stronger relationship earlier in the sample? How do different predictors relate to each other and interact in their relationships with attendance?


# II. Exploratory Data Analysis
## A. Exploring the Variables

### 1. Quantiles

```{r quantile ranges}
quan.att = quantile(attendance.data$Attendance, c(0.25, 0.5, 0.75), type = 1, na.rm = TRUE)
quan.fill = quantile(attendance.data$Fill.Rate, c(0.25, 0.5, 0.75), type = 1, na.rm = TRUE)
quan.cap = quantile(attendance.data$Stadium.Capacity, c(0.25, 0.5, 0.75), type = 1, na.rm = TRUE)
quan.att
quan.cap
quan.fill
```

The median for attendance is `r quan.att[2]`, with 25% and 75% quantiles of `r quan.att[1]` and `r quan.att[3]`. For stadium capacity, we have a median of `r quan.cap[2]` and quantiles of `r quan.cap[1]` and `r quan.cap[3]`. Fill rate has a median of `r quan.fill[2]`, a 25% quantile of `r quan.fill[1]`, and a 75% quantile of `r quan.fill[3]`.

```{r bad-data-removal}
attendance.data = attendance.data[-c(5436,5443,5450,5457, 4522, 4524, 4526, 4528, 4530, 4532),]
```

```{r incorrect-data}
attendance.data[4424, 24] = 12

attendance.data$Conference[which(attendance.data$Team=="Oregon State")] = "Pac-12"

attendance.data$Conference[which(attendance.data$Team=="USC")] = "Pac-12"

rutgers = attendance.data[which(attendance.data$Team=="Rutgers"),]
yr = 2000
while(yr<2013){
        rutgers$Conference[which(rutgers$Year==yr)] = "Big East"
        yr = yr+1
}
attendance.data$Conference[which(attendance.data$Team=="Rutgers")] = rutgers$Conference

boston_college = attendance.data[which(attendance.data$Team=="Boston College"),]
yr = 2000
while(yr<2005){
        boston_college$Conference[which(boston_college$Year==yr)] = "Big East"
        yr = yr+1
}
attendance.data$Conference[which(attendance.data$Team=="Boston College")] = boston_college$Conference

smu = attendance.data[which(attendance.data$Team=="SMU"),]
yr = 2005
while(yr<2013){
        smu$Conference[which(smu$Year==yr)] = "CUSA"
        yr = yr+1
}
attendance.data$Conference[which(attendance.data$Team=="SMU")] = smu$Conference

south_alabama = attendance.data[which(attendance.data$Team=="South Alabama"),]
yr = 2013
while(yr<2019){
        south_alabama$Conference[which(south_alabama$Year==yr)] = "Sun Belt"
        yr = yr+1
}
attendance.data$Conference[which(attendance.data$Team=="South Alabama")] = south_alabama$Conference

west_virginia = attendance.data[which(attendance.data$Team=="West Virginia"),]
yr = 2000
while(yr<2012){
        west_virginia$Conference[which(west_virginia$Year==yr)] = "Big East"
        yr = yr+1
}
attendance.data$Conference[which(attendance.data$Team=="West Virginia")] = west_virginia$Conference

memphis = attendance.data[which(attendance.data$Team=="Memphis"),]
memphis$Conference[which(memphis$Year==2013)] = "AAC"
attendance.data$Conference[which(attendance.data$Team=="Memphis")] = memphis$Conference

mid_tennessee = attendance.data[which(attendance.data$Team=="Middle Tennessee"),]
mid_tennessee$Conference[which(mid_tennessee$Year==2001)] = "Sun Belt"
attendance.data$Conference[which(attendance.data$Team=="Middle Tennessee")] = mid_tennessee$Conference

byu = attendance.data[which(attendance.data$Team=="BYU"),]
byu$Conference[which(byu$Year==2011)] = "Independent"
attendance.data$Conference[which(attendance.data$Team=="BYU")] = byu$Conference

fiu = attendance.data[which(attendance.data$Team=="FIU"),]
yr = 2002
while(yr<2005){
        fiu$Conference[which(fiu$Year==yr)] = "FCS"
        yr = yr+1
}
attendance.data$Conference[which(attendance.data$Team=="FIU")] = fiu$Conference

georgia_st = attendance.data[which(attendance.data$Team=="Georgia State"),]
yr = 2011
while(yr<2013){
        georgia_st$Conference[which(georgia_st$Year==yr)] = "FCS"
        yr = yr+1
}
attendance.data$Conference[which(attendance.data$Team=="Georgia State")] = georgia_st$Conference
```

